Spin canonical divisors of log stacky curves

### 2015-06-27 stacky_write_up/spin_cring.tex
section 3 transitions

\section{Inductive Lemmata}
\label{sec:induction}
First we present several lemmata which play a crucial role in the inductive step of the proof of the main theorem, Theorem ~\ref{thm:main}. In Subsection ~\ref{ssec:add-points} we prove Lemma ~\ref{lem:sat-1}, ~\ref{lem:sat-2}, and ~\ref{lem:sat-3} which describe the generators and relations of $R_{D + \frac{\alpha}{\beta}P},$ given those in $R_D,$ where $X$ is a curve, $D \in \di X \otimes \BQ$ and  $\frac{\alpha}{\beta} \in \BQ.$ Next, in Subsection ~\ref{ssec:raise-orders}, we prove Lemma ~\ref{lem:raise-stacky-order}, which the generators and relations of the log spin canonical ring curve with signature $(g; e_1', \ldots, e_\tau'; 0)$ to that of the log spin canonical ring of a curve with a certain subset of the $e_i'$ incremented by $2$. While Lemma ~\ref{lem:sat-1} may be applied in a fairly general context, Lemma ~\ref{lem:raise-stacky-orders} is more specific to half-canonical divisors, although even Lemma ~\ref{lem:raise-stacky-orders} has certain generalizations, as noted in Remark ~\ref{rem:raise-fractional-order}

@@

\ssec{Inductive Lemmata: Adding Points}
\label{ssec:add-points}
In this section, we prove various inductive statements under the following setup:
Let $\halfcan' \in \di \BQ \otimes X$ be $\BQ$-divisor of $X$ (and often a log half-canonical divisor of some $\sx'$), satisfying certain conditions about $\deg(\lfloor d \halfcan'\rfloor)$ for various values of $d\in\mathbb{N}$ and certain saturation conditions. Let $\halfcan=\halfcan'+P$ for some $P$ possibly in the support of $\halfcan$.  Then we will explicitly describe elements that generate $R_\halfcan$ over $\halfcan$' and initial terms of relations that generate $\initial(I')$ over $\initial(I)$ (allowing the construction of a corresponding Gr\"{o}bner Basis). \todo{Aaron: What new ideas does this description add that wasn't already described in the previous paragraph? Also, in general it's a good idea to be specific about what the most important lemmas are in the section, although I also added this to the general description of section 3.}

First, we introduce some notation from O'Dorney.  \cite{dorney:canonical}
\begin{defn}\label{den:lower-approximation}
If $\frac{\alpha}{\beta}\in \BQ$ (for $\alpha, \beta \in \mathbb{Z}, \beta> 0$ written in reduced form), then a \begin{bf}{best lower approximation}\end{bf} to $\frac{\alpha}{\beta}$ is a a fraction $\frac{c}{d}$ (written in reduced form with $d>0$) such that there is no $\frac{c'}{d'}$ with $0<d'<d$ and $\frac{c}{d}\le \frac{c'}{d'}$.  Moreover, the non-negative best lower approximations of $\frac{\alpha}{\beta}$ form a sequence
\[
	0=\frac{c_0}{d_0} < \frac{c_1}{d_1} < \ldots < \frac{c_r}{d_r} = \frac{\alpha}{\beta}.
\]
\todo{This Moreover statement shouldn't really be part of the definition.}
\end{defn}
\todo{add a picture of lattice}

In \cite{dorney:canonical}, O'Dorney uses best lower approximations to provide a minimal presentation for the canonical ring of a (not-necessarily effective) \todo{Can we remove this parenthetical? I don't see why the reader would assume effective.} divisor $D=\frac{\alpha_1}{\beta_1}P_1+\frac{\alpha_2}{\beta_2}P_2$ of $\BP_k^1$ \todo{This comes up a lot, but instead of saying $D$ is a divisor of $X$, I'd prefer to say $D \in \di X$ to be more precise. Of course, this is minor, and it's fine to keep it this way.} consisting of at most 2 points in its support, \todo{It might be clearer to say $\# \Supp(D) \leq 2.$} to show it is generated in degree $e=\max{\beta_1,\beta_2}$ with relations generated in degrees at most $2e$. \todo{Can you break the previous sentence into two sentences?} He \todo{Probably better not to say he, maybe give a theorem number where he does this?} further extends the result to more points, but only bounds the degree of minimal generation by a value on the order the least common multiple of the denominators of fractional points. \todo{This sentence seems very imprecise to me, and I wouldn't understand it if I hadn't read Evan's paper. Can you state this more precisely, or possibly just remove it.}

In Lemma \ref{lem:sat-1} we provide an inductive method to strengthen
O'Dorney's result \todo{cite theorem 8, Aaron: If you want to cite Theorem 8, just do \cite[Theorem 8]{dorney:canonical}} on $\BP^1_k$ to divisors $D = \sum_{i=1}^n \frac{\alpha_i
}{\beta_i} P_i$ for points $P_1, \ldots, P_n$ on a curve $X$ such that $H^
0(X, \lfloor D \rfloor) > 0$. Our result implies that $R_D$ is generated in
degrees at most $e = \max(\beta_i)$ with relations generated in
degrees at most $2e$. \todo{Are you sure you're not assuming $\frac{\alpha}{\beta}$ is effective? If not, doesn't what you say contradict Evan's example showing that his bounds of about lcm are asymptotically sharp?} Furthermore, given additional constraints on $H^0(
X, \lfloor dD \rfloor)$ for various values of $d$, we can apply the
same Riemann-Roch \todo{same as what? Maybe just say we can generalize his results to arbitrary genus?} arguments in arbitrary genus. 
To do this, we extend an inductive result \todo{Is ``inductive result'' a phrase? Can we just remove the word ``inductive'' every time it appears?} by Voight and
Zureick-Brown on stacky canonical rings to arbitrary $\BQ$-
divisors. \todo{cite theorem 8.3.1} \todo{Again, I don't think it can be arbitrary, and what are the $\BQ$ divisors on?}

First, however, we give a criteria to determine if a set of monomials generates the initial ideal of relations, and define notation for a type of monomial ordering. \todo{Perhaps say which parts of the proofs of which future lemmas it will be used in?}

\todo{Somehow, I feel like most of this page can be cut, it seems to mostly vaguely describe what we are doing, and takes up a lot of space.}

\begin{lem}\label{lem:relations_from_generators_induction} 
Suppose $\halfcan, \halfcan' \in \di X \otimes \BQ$ with $\halfcan=\halfcan'+\frac{\alpha}{\beta}P$, that $R_{\halfcan'}$ is generated by $x_1, \ldots, x_m,$
and $R_{\halfcan}$ is minimally generated by $y_1, \ldots, y_n$ over 
$R_{\halfcan'}$ with ideals of relations $I'$ and $I$ respectively \todo{Usually respectively is applied when two things are said at the same time, this doesn't seem to fit the usual gramattical pattern}.  
Suppose there are sets of monomials $S\subseteq R_\halfcan-R_{\halfcan'}$ and $T\subseteq R_\halfcan-(S\cup R_\halfcan')$, and an ordering $\prec$ such 
\begin{enumerate}
\item[(\ref{lem:relations_from_generators_induction}.1)] \todo{It would probably be clearer to just call these conditions 1,2,3 instead of 3.2.1...}  $T\prec S\prec k[x_1, \ldots, x_n]$ (meaning all monomials in $T$ precede all monomials in $S$ which proceed all monomials in $k[x_1, \ldots, x_n]$)
\item[(\ref{lem:relations_from_generators_induction}.2)] $S$ forms a $k$-basis for $R_\halfcan$ over $R_\halfcan'$
\item[(\ref{lem:relations_from_generators_induction}.3)] All monomials in $k[x_1, \ldots, x_m, y_1, \ldots, y_n]$ lie in 
\begin{align*}
	S	\cup \; \langle T\rangle \cup \; \initial_\prec(I') k[x_1, \ldots, x_m, y_1, \ldots, y_n]
\end{align*}
\end{enumerate}
Then,
\begin{align*}
	\initial(I) & = \initial(I') k[x_1, \ldots, x_m, y_1, \ldots, y_n]
	+ \langle T \rangle.
\end{align*}
\end{lem}
\begin{proof}
Suppose a monomial $f\in T$ has a pole order $a$ at $P$. Then, since $S$ forms a $k$-basis of $R_{\halfcan}$ over $R_{\halfcan'}$, we can write $f = (\sum_{g\in S'} C_g g)+r$ for some finite subset $S'\subseteq S_{\deg(f)}$ and $r\in R_{\halfcan'}$.  This demonstrates that 
\begin{align*}
	\initial(I) & \supseteq \initial(I') k[x_1, \ldots, x_m, y_1, \ldots, y_n] + \langle T \rangle.
\end{align*}
\todo{Aaron: I don't understand why this holds}
Since every monomial of $k[x_1, \ldots, x_m, y_1, \ldots, y_n]$ is an element of either $S$, $\langle T\rangle$, or $\initial_\prec(I') k[x_1, \ldots, x_m, y_1, \ldots, y_n]$,
\begin{align*}
	\initial(I) & = \initial(I') k[x_1, \ldots, x_m, y_1, \ldots, y_n] + \langle T \rangle.
\end{align*}
\todo{Aaron: Why can't the initial term lie in S?}
\todo{It might be helpful to say where you use conditions 1 2 and 3.}
\end{proof}

As suggested by the previous lemma, \todo{Reference the lemma, instead of using previous, since things might get moved around} we will need an ordering on monomials, and the following definition will prove useful in the ensuing lemma. \todo{Again, reference this}
\begin{defn}
We equip the graded ring $k[x_1, \ldots x_m]$, with weight $a_i$ and pole order $-\ord_P(x_i)$ attached to $x_i$, \todo{attached to $x_i$ doesn't really sound like the correct terminology to me} the graded $P$ reverse lexicographic order {\bf{g-P-revlex}} \todo{This name seems a bit funny looking to me} if for $f=\prod_{i=1}^m {x_i}^{q_i}\ne g=\prod_{i=1}^m {x_i}^{r_i}$, $f\prec g$ iff \todo{It's better not to use abbreviations like iff in papers. Also, usually in definitions, one uses if in place of iff.} one of the following holds
\begin{enumerate}
\item $\deg(f) < \deg(g)$
\item $\deg(f) = \deg(g)$ and $-\ord(f) > -\ord(g)$ \todo{Order at which point, in this line and the next}
\item $\deg(f) = \deg(g)$, $-\ord(f) = -\ord(g)$, and for $i$ maximal such that $q_i-r_i$ is non-zero, that expression is negative. \todo{I do not understand what this means, maximal under what ordering? Also, }
\end{enumerate}
\end{defn}
We leave it as an exercise to verify that this is an order.  \todo{check for ourselves that this actually is, Aaron: I suggest not to use the phrase ``leave as an exercise'' in a paper}
We now offer a lemma that allows us to inductively add $\mathbb{Q}$-points to a divisor and describe the new generators and relations.

@@

\todo{Something about fractional weight modular forms?}

@@

\todo{transition!}

>>>>

\section{Inductive Lemmata}
\label{sec:induction}
First we present several lemmata which play a crucial role in the inductive step of the proof of the main theorem, Theorem ~\ref{thm:main}. In Subsection ~\ref{ssec:add-points} we prove Lemma ~\ref{lem:sat-1}, ~\ref{lem:sat-2}, and ~\ref{lem:sat-3} which describe the generators and relations of $R_{D + \frac{\alpha}{\beta}P},$ given those in $R_D,$ where $X$ is a curve, $D \in \di X \otimes \BQ$ and  $\frac{\alpha}{\beta} \in \BQ.$ Next, in Subsection ~\ref{ssec:raise-orders}, we prove Lemma ~\ref{lem:raise-stacky-order}, which the generators and relations of the log spin canonical ring curve with signature $(g; e_1', \ldots, e_\tau'; 0)$ to that of the log spin canonical ring of a curve with a certain subset of the $e_i'$ incremented by $2$. While Lemma ~\ref{lem:sat-1} may be applied in a fairly general context, Lemma ~\ref{lem:sat-2} and ~\ref{lem:sat-3} are more specific to half-canonical divisors.

@@

\ssec{Inductive Lemmata: Adding Points}
\label{ssec:add-points}
In this section, we prove various inductive statements under the following setup:
Let $\halfcan'$ be $\BQ$-divisor of $X$ (and often a log half-canonical divisor of some $\sx'$), satisfying certain conditions about $\deg(\lfloor d \halfcan'\rfloor)$ for various values of $d\in\mathbb{N}$ and certain saturation conditions. Let $\halfcan=\halfcan'+P$ for some $P$ possibly in the support of $\halfcan$.  Then we will explicitly describe elements that generate $R_\halfcan$ over $\halfcan$' and initial terms of relations that generate $\initial(I')$ over $\initial(I)$ (allowing the construction of a corresponding Gr\"{o}bner Basis).

First, let's introduce some notation from O'Dorney.  \cite{dorney:canonical}
\begin{defn}\label{den:lower-approximation}
If $\frac{\alpha}{\beta}\in \BQ$ (for $\alpha\in \mathbb{Z}, \beta\in \mathbb{N}$ written in reduced form), then a \begin{bf}{best lower approximation}\end{bf} to $\frac{\alpha}{\beta}$ is a a fraction $\frac{c}{d}$ (written in reduced form with $d>0$) such that there is no $\frac{c'}{d'}$ with $0<d'<d$ and $\frac{c}{d}\le \frac{c'}{d'}$.  Moreover, the non-negative best lower approximations of $\frac{\alpha}{\beta}$ form a sequence
\[
	0=\frac{c_0}{d_0} < \frac{c_1}{d_1} < \ldots < \frac{c_r}{d_r} = \frac{\alpha}{\beta}.
\]
\end{defn}
\todo{add a picture of lattice}

In ~\cite{dorney:canonical}, O'Dorney uses best lower approximations to provide a minimal presentation for the canonical ring of a (not-necessarily effective) divisor $D=\frac{\alpha_1}{\beta_1}P_1+\frac{\alpha_2}{\beta_2}P_2$ of $\mathbb{P}^1$ consisting of at most 2 points in its support, showing that it is generated in degree $e=\max{\beta_1,\beta_2}$ with relations generated in degrees at most $2e$.  He further extends the result to more points, but only bounds the degree of minimal generation by a value on the order the least common multiple of the denominators of fractional points.

In Lemma ~\ref{lem:sat-1} we provide an inductive method to strengthen
O'Dorney's result \cite[Theorem 8]{dorney:canonical} on $\BP^1_k$ to divisors $D = \sum_{i=1}^n \frac{\alpha_i
}{\beta_i} P_i$ for points $P_1, \ldots, P_n$ on a curve $X$ such that $H^
0(X, \lfloor D \rfloor) > 0$; our result implies that $R_D$ is generated in
degrees at most $e = \max(\beta_i)$ with relations generated in
degrees at most $2e$.  Furthermore, given additional constraints on $H^0(
X, \lfloor dD \rfloor)$ for various values of $d$, we can apply the
same Riemann-Roch arguments in arbitrary genus. 
To do this, we extend an inductive result by Voight and
Zureick-Brown ~\cite[Theorem 8.3.1]{vzb:stacky} on stacky canonical rings to arbitrary $\BQ$-
divisors.

First, however, we give a criteria to determine if a set of monomials generates the initial ideal of relations, and define notation for a type of monomial ordering.

\begin{lem}\label{lem:relations_from_generators_induction} 
Suppose $\halfcan$ and $\halfcan'$ are $\BQ$ divisors on a curve $X$ with $\halfcan=\halfcan'+\frac{\alpha}{\beta}P$, that $R_{\halfcan'}$ is generated by $x_1, \ldots, x_m$ and $R_{\halfcan}$ is minimally generated by $y_1, \ldots, y_n$ over $R_{\halfcan'}$ with ideals of relations $I'$ and $I$ respectively.  Suppose there are sets of monomials $S\subseteq R_\halfcan-R_{\halfcan'}$ and $T\subseteq R_\halfcan-(S\cup R_\halfcan')$, and an ordering $\prec$ such 
\begin{enumerate}
\item[(\ref{lem:relations_from_generators_induction}.1)]  $T\prec S\prec k[x_1, \ldots, x_n]$ (meaning all monomials in $T$ precede all monomials in $S$ which proceed all monomials in $k[x_1, \ldots, x_n]$)
\item[(\ref{lem:relations_from_generators_induction}.2)] $S$ forms a $k$-basis for $R_\halfcan$ over $R_\halfcan'$
\item[(\ref{lem:relations_from_generators_induction}.3)] All monomials in $k[x_1, \ldots, x_m, y_1, \ldots, y_n]$ lie in 
\begin{align*}
	& S \\
	\cup \; &\langle T\rangle \\
	\cup \; & \initial_\prec{I'} k[x_1, \ldots, x_m, y_1, \ldots, y_n]
\end{align*}
\end{enumerate}
Then 
\begin{align*}
	\initial(I) & = \initial(I') k[x_1, \ldots, x_m, y_1, \ldots, y_n]\\
		     & + \langle T \rangle.
\end{align*}
\end{lem}
\begin{proof}
Suppose monomial $f\in T$ has a pole order $a$ at $P$.  Then since $S$ forms a $k$-basis of $R_{\halfcan}$ over $R_{\halfcan'}$, we can write $f + (\sum_{g\in S'} C_g g)+r$ for some finite subset $S'\subseteq S_{\deg(f)}$ and $r\in R_{\halfcan'}$.  This demonstrates that 
\begin{align*}
	\initial(I) & \supseteq \initial(I') k[x_1, \ldots, x_m, y_1, \ldots, y_n]\\
		     & + \langle T \rangle.
\end{align*}
Since every monomial of $k[x_1, \ldots, x_m, y_1, \ldots, y_n]$ is an element of either $S$, $\langle T\rangle$, or $\initial_\prec(I') k[x_1, \ldots, x_m, y_1, \ldots, y_n]$,
\begin{align*}
	\initial(I) & = \initial(I') k[x_1, \ldots, x_m, y_1, \ldots, y_n]\\
		     & + \langle T \rangle.
\end{align*}
\end{proof}

As suggested by the previous lemma, we will need an ordering on monomials, and the following definition will prove useful in the ensuing lemma.
\begin{defn}
We equip the graded ring $k[x_1, \ldots x_m]$, with weight $a_i$ and pole order $-\ord_P(x_i)$ attached to $x_i$, the graded $P$ reverse lexicographic order {\bf{g-P-revlex}} if for $f=\prod_{i=1}^m {x_i}^{q_i}\ne g=\prod_{i=1}^m {x_i}^{r_i}$, $f\prec g$ iff one of the following holds
\begin{enumerate}
\item $\deg(f) < \deg(g)$
\item $\deg(f) = \deg(g)$ and $-\ord(f) > -\ord(g)$
\item $\deg(f) = \deg(g)$, $-\ord(f) = -\ord(g)$, and for $i$ maximal such that $q_i-r_i$ is non-zero, that expression is negative.  
\end{enumerate}
\end{defn}
We leave it as an exercise to verify that this is an order.  \todo{check for ourselves that this actually is, Aaron: I suggest not to use the phrase ``leave as an exercise'' in a paper}
We now offer a lemma that allows us to inductively add $\mathbb{Q}$-points to a divisor and describe the new generators and relations.

@@

Lemma \ref{lem:sat-1} leads to the the following corollaries.
\todo{do we want this here?}
\begin{cor}\label{cor-gen-0-with-log}
If $(\sx, \halfcan, \Delta)$ is a stacky curve with signature $(g; e_1, \ldots, e_n; \delta)$ with $g=0$ and $\delta>0$, then $R_\halfcan$ is generated in degree at most $e=\max(e_i)$ with relations generated in degree at most $2e$.
\end{cor}
\begin{proof}
By remark \ref{rem:delta-not-1}, $\delta\ge 2$ so $\halfcan$ is linearly equivalent to an effective divisor.  Thus, we can recursively apply Lemma \ref{lem:sat-1} to prove the corollary.
\end{proof}

\todo{Something about fractional weight modular forms?}
\iffalse{
\begin{cor}
Let $R_\halfcan$ be the fractional weight canonical ring associated to a fractional $\frac{2\alpha}{\beta}$-weight cusp forms with $\frac{\alpha}{\beta}$ in reduced form and $\beta\ge 3$, with signature $(g; e_1, \ldots, e_n; \delta)$ where $g=0$.  Then $R_\halfcan$ is generated in degrees at most $e = \max(e_i)$ with relations generated in degrees at most $2e$.
\end{cor}
\begin{proof}
Let $K$ be the log canonical divisor of the curve $(\sx, \Delta)$ associated to the $\frac{2\alpha}{\beta}$ weight cusp forms.  We see that $\frac{\beta}{\alpha}\halfcan_\sx\sim K_\sx$, so the floors of each

Then $\beta \halfcan_\sx \sim \alpha K_{\sx}$ so $\beta\lfloor so $\beta | \alpha \deg(K_\sx)$; since $\gcd(\alpha, \beta)=1$, $\beta| \deg(K_\sx)\ge -2$.  Thus, since $\beta\ge 3$, $\deg(K_\sx)\ge 0$.  Thus $\deg(
\end{proof}
}\fi

We now restrict our attention to log canonical rings of stacky curves.  The previous lemma deals with most relevant induction cases when inducting from a half-canonical ring saturated at one.  We complement it with the following two lemmas that allow us to induct under certain conditions when the half-canonical ring is saturated at two and three.

@@

One can prove similar results in cases with different conditions on saturation, base-point freeness, and the coefficients of added points, but only these three cases (Lemmas ~\ref{lem:sat-1}, ~\ref{lem:sat-2}, and ~\ref{lem:sat-3}) are needed for the remainder of this paper.  Instead we turn now to inductive methods to incrementally increase the $e_i$s.  Combined with the previous lemma on adding points, these will provide all the necessary inductive tools.
